---
title: 'Earnings Manipulation - Benford\'s Law'
author: "Lu Chen"
output:
  html_document: default
  pdf_document: default
Indent: false
---

```{r}
setwd("")
df1n <- read.csv('cleaned_data.csv',header = TRUE)
attach(data)
library(ggplot2)
library(dplyr)
library(data.table)
library(benford.analysis)
library(plm)
library(knitr)
library(kableExtra)
library(zoo)
```

**1.** Plot the distribution of the first digit
```{r fig1, fig.height = 3, fig.width = 5, fig.align = "center"}
revenue_bf <- data.frame(benford(abs(df1n$revt), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3)$bfd)[,c("digits","data.dist", "benford.dist", "data.second.order.dist")]
cash_bf <- data.frame(benford(abs(df1n$oancf), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3)$bfd)[,c("digits","data.dist", "benford.dist","data.second.order.dist")]
asset_bf <- data.frame(benford(abs(df1n$at), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3)$bfd)[,c("digits","data.dist", "benford.dist","data.second.order.dist")]

colnames(revenue_bf) <- c('digits','revenue', 'benford_revenue','second_revenue')
colnames(cash_bf) <- c('digits','cashflow', 'benford_cashflow','second_cashflow')
colnames(asset_bf) <- c('digits','totalasset', 'benford_at','second_at')

digits_df <- merge(merge(revenue_bf, cash_bf), asset_bf)

# plotting revenue and benford revenue
p1=ggplot(digits_df) +
  geom_line(aes(y=digits, x=revenue, colour="revenue"))+
  geom_line(aes(y=digits, x = benford_revenue, colour="benford_revenue"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("First digit distribution of actual revnue and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("revenue" = 'red', "benford_revenue" = 'blue'))+
  theme_classic()

# plotting cash flow and benford cash flow
p2=ggplot(digits_df) +
  geom_line(aes(y=digits, x=cashflow, colour="cashflow"))+
  geom_line(aes(y=digits, x = benford_cashflow, colour="benford_cashflow"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("First digit distribution of actual asset and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("cashflow" = 'red', "benford_cashflow" = 'blue'))+
  theme_classic()

# plotting asset and benford asset
p3=ggplot(digits_df) +
  geom_line(aes(y=digits, x=totalasset, colour="asset"))+
  geom_line(aes(y=digits, x = benford_revenue, colour="benford_asset"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("First digit distribution of actual asset and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("asset" = 'red', "benford_asset" = 'blue')) + 
  theme_classic()

p4=ggplot(digits_df) +
  geom_line(aes(digits, revenue, col='revenue'), linetype=1)+
  geom_line(aes(digits, cashflow, col='cash flow'), linetype=2)+
  geom_line(aes(digits, totalasset, col='total asset'), linetype=4)+
  #scale_x_continuous(limits=c(1985, 2017),breaks=seq(1985, 2017, 5))+
  scale_color_manual(name="", values=c("revenue"="darkgreen","cash flow"="blue",'total asset'='red'))+
  labs(x = 'digits', y = 'frequency')+
  ggtitle("Distribution Comparison of first-digit asset, revenue and cashflow")+
  theme_classic()

library("Rmisc")
library("plyr")
multiplot(p1,p2,p3,p4)
```

**2.** Report the χ2 test statistic and interpret its significance. 
```{r}
c_revenue <- chisq(benford(abs(df1n$revt), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$statistic['X-squared']
c_cash <- chisq(benford(abs(df1n$oancf), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$statistic['X-squared']
c_asset <- chisq(benford(abs(df1n$at), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$statistic['X-squared']

p_revenue <- chisq(benford(abs(df1n$revt), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$p.value
p_cash <- chisq(benford(abs(df1n$oancf), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$p.value
p_asset <- chisq(benford(abs(df1n$at), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$p.value

chi <- data.frame(c_revenue, c_cash, c_asset)
p_value <- data.frame(p_revenue, p_cash, p_asset)

chi1 <- t(chi)
p_value1 <- t(p_value)
colnames(p_value1) <- 'p_value'
chi_test <- cbind(chi1, p_value1)
kable(chi_test, format = "markdown")
```

**3.** Repeat for second digit
```{r fig3, fig.height = 3, fig.width = 5, fig.align = "center"}
# plotting revenue and benford revenue
p5=ggplot(digits_df) +
  geom_line(aes(y=digits, x=second_revenue, colour="revenue"))+
  geom_line(aes(y=digits, x = benford_revenue, colour="benford_revenue"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("Second digit distribution of actual revnue and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("revenue" = 'red', "benford_revenue" = 'blue'))+
  theme_classic()

# plotting cash flow and benford cash flow
p6=ggplot(digits_df) +
  geom_line(aes(y=digits, x=second_cashflow, colour="cashflow"))+
  geom_line(aes(y=digits, x = benford_cashflow, colour="benford_cashflow"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("Second digit distribution of actual asset and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("cashflow" = 'red', "benford_cashflow" = 'blue'))+
  theme_classic()

# plotting asset and benford asset
p7=ggplot(digits_df) +
  geom_line(aes(y=digits, x=second_at, colour="asset"))+
  geom_line(aes(y=digits, x = benford_revenue, colour="benford_asset"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("Second digit distribution of actual asset and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("asset" = 'red', "benford_asset" = 'blue'))+
  theme_classic()

p8=ggplot(digits_df) +
  geom_line(aes(digits, second_revenue, col='revenue'), linetype=1)+
  geom_line(aes(digits, second_cashflow, col='cash flow'), linetype=2)+
  geom_line(aes(digits, second_at, col='total asset'), linetype=4)+
  scale_color_manual(name="", values=c("revenue"="darkgreen","cash flow"="blue",'total asset'='red'))+
  labs(x = 'digits', y = 'frequency')+
  theme_classic()+
  ggtitle("Distribution Comparison of second-digit asset, revenue and cashflow")

library("Rmisc")
library("plyr")
multiplot(p5,p6,p7,p8)

```

**5.**Summrize: for both first and second digit, we use three variables(total asset, cash flow and revenue) to compare the distribution of actual data with therotical distribution - benford. It turns out that for first-digit, neither variable deviate from the theoretical pattern. However, for second-digit, all three variables, assets, revenue, and operating cash flows seem to deviate from Benford’s law. Note that even operating cash flows can be manipulated if firms are channel stuffing.
