---
title: 'Earnings Manipulation - Benford\'s Law'
author: "Lu Chen"
output:
  html_document: default
  pdf_document: default
Indent: false
---

```{r}
setwd("/Users/chenlu/Google Drive/Summer/6030 Accounting/data assignment1")
df1n <- read.csv('cleaned_data.csv',header = TRUE)
attach(data)
library(ggplot2)
library(dplyr)
library(data.table)
library(benford.analysis)
library(plm)
library(knitr)
library(kableExtra)
library(zoo)
```


## 1. Benford Analysis

**1.** Plot the distribution of the first digit
```{r fig1, fig.height = 3, fig.width = 5, fig.align = "center"}
revenue_bf <- data.frame(benford(abs(df1n$revt), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3)$bfd)[,c("digits","data.dist", "benford.dist", "data.second.order.dist")]
cash_bf <- data.frame(benford(abs(df1n$oancf), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3)$bfd)[,c("digits","data.dist", "benford.dist","data.second.order.dist")]
asset_bf <- data.frame(benford(abs(df1n$at), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3)$bfd)[,c("digits","data.dist", "benford.dist","data.second.order.dist")]

colnames(revenue_bf) <- c('digits','revenue', 'benford_revenue','second_revenue')
colnames(cash_bf) <- c('digits','cashflow', 'benford_cashflow','second_cashflow')
colnames(asset_bf) <- c('digits','totalasset', 'benford_at','second_at')

digits_df <- merge(merge(revenue_bf, cash_bf), asset_bf)

# plotting revenue and benford revenue
p1=ggplot(digits_df) +
  geom_line(aes(y=digits, x=revenue, colour="revenue"))+
  geom_line(aes(y=digits, x = benford_revenue, colour="benford_revenue"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("First digit distribution of actual revnue and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("revenue" = 'red', "benford_revenue" = 'blue'))+
  theme_classic()

# plotting cash flow and benford cash flow
p2=ggplot(digits_df) +
  geom_line(aes(y=digits, x=cashflow, colour="cashflow"))+
  geom_line(aes(y=digits, x = benford_cashflow, colour="benford_cashflow"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("First digit distribution of actual asset and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("cashflow" = 'red', "benford_cashflow" = 'blue'))+
  theme_classic()

# plotting asset and benford asset
p3=ggplot(digits_df) +
  geom_line(aes(y=digits, x=totalasset, colour="asset"))+
  geom_line(aes(y=digits, x = benford_revenue, colour="benford_asset"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("First digit distribution of actual asset and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("asset" = 'red', "benford_asset" = 'blue')) + 
  theme_classic()

p4=ggplot(digits_df) +
  geom_line(aes(digits, revenue, col='revenue'), linetype=1)+
  geom_line(aes(digits, cashflow, col='cash flow'), linetype=2)+
  geom_line(aes(digits, totalasset, col='total asset'), linetype=4)+
  #scale_x_continuous(limits=c(1985, 2017),breaks=seq(1985, 2017, 5))+
  scale_color_manual(name="", values=c("revenue"="darkgreen","cash flow"="blue",'total asset'='red'))+
  labs(x = 'digits', y = 'frequency')+
  ggtitle("Distribution Comparison of first-digit asset, revenue and cashflow")+
  theme_classic()

library("Rmisc")
library("plyr")
multiplot(p1,p2,p3,p4)
```

**2.** Report the χ2 test statistic and interpret its significance. 
```{r}
c_revenue <- chisq(benford(abs(df1n$revt), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$statistic['X-squared']
c_cash <- chisq(benford(abs(df1n$oancf), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$statistic['X-squared']
c_asset <- chisq(benford(abs(df1n$at), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$statistic['X-squared']

p_revenue <- chisq(benford(abs(df1n$revt), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$p.value
p_cash <- chisq(benford(abs(df1n$oancf), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$p.value
p_asset <- chisq(benford(abs(df1n$at), number.of.digits = 1, sign = 'positive', discrete = TRUE, round = 3))$p.value

chi <- data.frame(c_revenue, c_cash, c_asset)
p_value <- data.frame(p_revenue, p_cash, p_asset)

chi1 <- t(chi)
p_value1 <- t(p_value)
colnames(p_value1) <- 'p_value'
chi_test <- cbind(chi1, p_value1)
kable(chi_test, format = "markdown")
```

**3.** Repeat for second digit
```{r fig3, fig.height = 3, fig.width = 5, fig.align = "center"}
# plotting revenue and benford revenue
p5=ggplot(digits_df) +
  geom_line(aes(y=digits, x=second_revenue, colour="revenue"))+
  geom_line(aes(y=digits, x = benford_revenue, colour="benford_revenue"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("Second digit distribution of actual revnue and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("revenue" = 'red', "benford_revenue" = 'blue'))+
  theme_classic()

# plotting cash flow and benford cash flow
p6=ggplot(digits_df) +
  geom_line(aes(y=digits, x=second_cashflow, colour="cashflow"))+
  geom_line(aes(y=digits, x = benford_cashflow, colour="benford_cashflow"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("Second digit distribution of actual asset and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("cashflow" = 'red', "benford_cashflow" = 'blue'))+
  theme_classic()

# plotting asset and benford asset
p7=ggplot(digits_df) +
  geom_line(aes(y=digits, x=second_at, colour="asset"))+
  geom_line(aes(y=digits, x = benford_revenue, colour="benford_asset"),linetype = 2) +
  labs(x = 'digits', y = 'frequency') +
  ggtitle("Second digit distribution of actual asset and theoretical benford ")+
  scale_color_manual(name = "",
                     values = c("asset" = 'red', "benford_asset" = 'blue'))+
  theme_classic()

p8=ggplot(digits_df) +
  geom_line(aes(digits, second_revenue, col='revenue'), linetype=1)+
  geom_line(aes(digits, second_cashflow, col='cash flow'), linetype=2)+
  geom_line(aes(digits, second_at, col='total asset'), linetype=4)+
  scale_color_manual(name="", values=c("revenue"="darkgreen","cash flow"="blue",'total asset'='red'))+
  labs(x = 'digits', y = 'frequency')+
  theme_classic()+
  ggtitle("Distribution Comparison of second-digit asset, revenue and cashflow")

library("Rmisc")
library("plyr")
multiplot(p5,p6,p7,p8)

```

**4.** 
```{r fig3, fig.height = 3, fig.width = 5, fig.align = "center"}
roa_df <- df1n %>% 
  group_by(gvkey) %>% 
  mutate(at_lag1 = dplyr::lag(at, k = 1)) %>% 
  mutate(roa = ni / at_lag1) %>% 
  filter(roa >= -.1 & roa <= .1) %>% 
  mutate(roa_lag1 = dplyr::lag(roa, k=1),
         roa_change = roa - roa_lag1)

ggplot(subset(roa_df, roa_change != 'NA')) + 
  geom_histogram(aes(roa_change), bins = 30)+
  labs(x = 'ROA from year t to t-1 for each firm') + 
  geom_vline(xintercept=0, col='red', linetype=4)+
  theme_classic()
```
### Comparing ROA and EPS changing trend:EPS is more asymmetric and in the direction of positive side, indicating firms are reported to perform better YoY earnings per share than return on asset. Both performances asymmetric around 0.
### Possible explanations are: 1) earnings management 2) firms working harder and 3) inflation or other general trends that make small improvements more likely than small set backs.
### EPS equals to net income divided by total number of issued stock shares; ROA equals to net income deivided by lagged total assets. From the calculation, we can learn that they have the same numerator but different denominator. Firm have the ability to manage the number of shares outstanding (EPS denominator) by issuing more shares or repurchasing them back. Also if the firm assume the investors look at EPS more often than ROA, they have the incentive to drive EPS harder. 


